Cross-scale integration of data and knowledge for predicting species ranges

### 2014-08-19 tex/manuscript.tex
Added last remaining comments

% available colors:
% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!40]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

\title{Cross-scale integration of data and knowledge for predicting species ranges}
\date{}
\author[1,2]{Matthew V. Talluto}
\author[1,2]{Isabelle Boulangeat}
\author[3]{Aitor Ameztegui}
\author[4]{Isabelle Aubin}
\author[1,2,5]{Dominique Berteaux}
\author[1,2]{Alyssa Butler}
%\author[10]{Fred Doyon} %??
\author[6]{C. Ronnie Drever}
\author[7]{Marie-Josée Fortin}
\author[1]{Tony Franceschini}
\author[8]{Jean Liénard}
\author[4]{Dan McKenney}
\author[2,3]{Kevin A. Solarik}
\author[8]{Nick Strigul}
\author[9, 10]{Wilfried Thuiller}
\author[1,2]{Dominique Gravel}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[7]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[8]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
\affil[9]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[10]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}

@@

>>>>

% available colors:
% red, green, blue, cyan, magenta, yellow, gray, drakfray, lightgray, brown, lime, olive, orange, pink, purple, teal, violet
%
\newcommand{\mt}[1]{\todo[color=blue!40]{\textsuperscript{MT}#1}}
\newcommand{\ib}[1]{\todo[color=magenta!40]{\textsuperscript{IB}#1}} 				
\renewcommand{\aa}[1]{\todo[color=violet!40]{\textsuperscript{Aitor A}#1}}
\newcommand{\ia}[1]{\todo[color=blue!20]{\textsuperscript{Isabelle A}#1}}
\newcommand{\db}[1]{\todo[color=violet!20]{\textsuperscript{Dominique B}#1}}
\newcommand{\ab}[1]{\todo[color=teal!40]{\textsuperscript{Alyssa B}#1}}
\newcommand{\fd}[1]{\todo[color=orange!30]{\textsuperscript{Frédérik D}#1}}
%\newcommand{\rd}[1]{\todo[color=orange!40]{\textsuperscript{Ronnie D}#1}}
\newcommand{\mjf}[1]{\todo[color=green!60]{\textsuperscript{Marie-Josée F.}#1}}
\newcommand{\tf}[1]{\todo[color=yellow!60]{\textsuperscript{Tony F.}#1}}
\newcommand{\dm}[1]{\todo[color=lime!50]{\textsuperscript{Dan M.}#1}}
%\newcommand{\ks}[1]{\todo[color=green!40]{\textsuperscript{Kevin S}#1}}
\newcommand{\ns}[1]{\todo[backgroundcolor=red!100, linecolor=orange!100, bordercolor=cyan!100]{{\color{magenta!40}\textsuperscript{Strigul}#1}}}
\newcommand{\wt}[1]{\todo[color=orange!80]{\textsuperscript{Wilfried T}#1}}
\newcommand{\dg}[1]{\todo[color=pink!40]{\textsuperscript{DG}#1}}

@@

\title{Cross-scale integration of data and knowledge for predicting species ranges}
\date{}
\author[1,2]{Matthew V. Talluto}
\author[1,2]{Isabelle Boulangeat}
\author[3]{Aitor Ameztegui}
\author[4]{Isabelle Aubin}
\author[1,2,5]{Dominique Berteaux}
\author[1,2]{Alyssa Butler}
\author[6,7]{Frédérik Doyon} %??
\author[8]{C. Ronnie Drever}
\author[9]{Marie-Josée Fortin}
\author[1]{Tony Franceschini}
\author[10]{Jean Liénard}
\author[4]{Dan McKenney}
\author[2,3]{Kevin A. Solarik}
\author[10]{Nickolay Strigul}
\author[11, 12]{Wilfried Thuiller}
\author[1,2]{Dominique Gravel}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Quebec Centre for Biodiversity Science, Montreal, Quebec, Canada}
\affil[3]{Centre d'Étude de la Forêt, Département des sciences biologiques, Université du Québec à Montréal, Montreal, Quebec, Canada}
\affil[4]{Great Lakes Forestry Centre, Canadian Forest Service, Natural Resources Canada, Sault Ste Marie, Ontario, Canada}
\affil[5]{Centre for Northern Studies, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[6]{Université du Québec en Outaouais, Gatineau, Quebec, Canada}
\affil[7]{Institut des Sciences de la Forêt Tempérée (ISFORT), Ripon, Quebec, Canada}
\affil[8]{The Nature Conservancy Canada, Ottawa, Ontario, Canada}
\affil[9]{Department of Ecology and Evolutionary Biology, University of Toronto, Toronto, Ontario, Canada}
\affil[10]{Department of Mathematics, Washington State University Vancouver, Washington, USA}
\affil[11]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[12]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}

@@

\section*{Acknowledgements}
We acknowledge funding from the Quebec Centre for Biodiversity Science as well as NSERC strategic grant 430393-12. Comments from Charles Canham and William Godsoe improved a previous version of this manuscript.

### 2014-08-19 tex/model_description.tex
Added last remaining comments

The role of the metamodel model is to integrate data at the same ecological scale of predictions.
Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
Formally, we will add an additional model, \(\theta_S\), that is based on a different set of hypotheses and that makes predictions \((\psi_S)\) based on an additional dataset \((X_S, D_S)\):
%-----
\begin{equation}
\label{eq:model2-1}
	\psi_S = g(\theta_S, X_S, D_S)
\end{equation}
%-----

For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}. 
In many cases, computing this likelihood will be challenging.
The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
We simply use the posterior distribution of $\theta_M$ from the previous step (Eq. \ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the likelihood of $\theta_M$ in light of the new data and the prior knowledge, as before:
%-----
\begin{equation}
\label{eq:integrated2}
	\overbrace{p( \theta_M \mid X_M, D_M, \theta_S, X_S, D_S )}^\text{integrated posterior}
	\propto 
	\overbrace{p( \theta_S \mid \theta_M, X_S, D_S )}^\text{likelihood}
	\overbrace{p( X_M \mid D_M, \theta_M ) P( \theta_M )}^{\text{prior for } \theta_M}
	\overbrace{p(\theta_S)}^{\text{prior for } \theta_S}
\end{equation}
%-----

>>>>

The role of the metamodel is to integrate data at the same ecological scale of predictions.
Additional sub-models require outputs that should be comparable to this given ecological scale (e.g. constraining presence or absence on the landscape).
Formally, we will add an additional model, \(\theta_S\), that is based on a different set of hypotheses and that makes predictions \((\psi_S)\) based on an additional dataset \((X_S, D_S)\):
%-----
\begin{equation}
\label{eq:model2-1}
	\psi_S = g(\theta_S, X_S, D_S)
\end{equation}
%-----

For integration to be successful, it must be possible to compute the likelihood of this prediction given the metamodel, \(p \left( \psi_S \mid \theta_M ) = \theta_S \mid \theta_M, X_S, D_S \right)\), thus, we refer to $g$ as the \emph{scaling function}. 
In many cases, computing this likelihood will be challenging.
The simplest case is when the parameters of \(\theta_S\) are directly comparable to those in \(\theta_M\) (see Example 1), or when the scaling is performed directly within the sub-model (see Example 2), but other solutions are possible.
This new likelihood, which integrates the information from the second model into the first, can be treated as new ``data'' to evaluate the parameters of \(\theta_M\).
We simply use the posterior distribution of $\theta_M$ from the previous step (Eq. \ref{eq:sdm4}) as the new prior probability of $\theta_M$, and evaluate the likelihood of $\theta_M$ in light of the new data and the prior knowledge, as before:
\fd{Shouldn't there be "Os" somewhere here?\\

like:   "We simply use the posterior distibution of Om from the previous step (eq.11) as the new prior probability of Os, and evaluate the likelihood of Om in the ..."}
%-----
\begin{equation}
\label{eq:integrated2}
	\overbrace{p( \theta_M \mid X_M, D_M, \theta_S, X_S, D_S )}^\text{integrated posterior}
	\propto 
	\overbrace{p( \theta_S \mid \theta_M, X_S, D_S )}^\text{likelihood}
	\overbrace{p( X_M \mid D_M, \theta_M ) P( \theta_M )}^{\text{prior for } \theta_M}
	\overbrace{p(\theta_S)}^{\text{prior for } \theta_S}
\end{equation}
%-----
